---
title: america & americans
author: ''
date: '2020-01-27'
slug: america-americans
categories: []
tags: []
---


```{r message=FALSE, warning=FALSE, include=FALSE}
tidycensus::census_api_key("b508704c99f3ae9bc5b5e7c41e3dd77e59d52722")
```

Identifying US states & counties with demographics (here, race & ethnicity) most similar to the US overall.  Via Kullback–Leibler divergence, or relative entropy.



## Using tidycensus

Race-ethnicity distinctions made here are mututally exclusive.  'White', 'Black', 'AIAN', 'Asian', 'NativeHawaiian', 'SomeOtherRace', and 'TwoOrMoreRaces' categories represent non-Hispanic populations.   American Community Survey, 2014-2018 5-year estimates. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)

variable <- c('DP05_0071P', 'DP05_0077P', 'DP05_0078P',
              'DP05_0079P', 'DP05_0080P', 'DP05_0081P',
              'DP05_0082P', 'DP05_0083P' )

variable_name <- c('Hispanic', 'White', 'Black',
                   'AIAN', 'Asian', 'NativeHawaiian',
                   'SomeOtherRace', 'TwoOrMoreRaces' )

gen <-  lapply(c('us', 'state', 'county'), function (x) {
  tidycensus::get_acs(geography = x,
                      variables = variable,
                      year = 2018,
                      survey = 'acs5') })   %>%
  bind_rows() %>%
  left_join(data.frame(variable, 
                       variable_name, 
                       stringsAsFactors = FALSE))%>%
  select(-variable, -moe) %>%
  rename(variable = variable_name) %>%
  select(GEOID:NAME, variable, estimate) %>%
  mutate(estimate = estimate/100)

usa <- gen %>% filter(GEOID == 1) %>% 
  rename(usa = estimate) %>% 
  select(variable, usa)
```




## Comparing probability distributions

Distance between probability distributions  --  Old hat thought: **Iowa & New Hampshire are very White**.  Not especially well representative of the US, and certainly not the Democratic Party.  Perhaps not the best locales for first-in-the-nation caucuses/primary.  

```{r message=FALSE, warning=FALSE}
gen %>%
  filter(GEOID %in% c('1', '35', '17', '19', '33') &
           !variable %in% c('NativeHawaiian', 
                            'SomeOtherRace', 
                            'TwoOrMoreRaces')) %>% #
  select(-GEOID) %>%
  spread(variable, estimate) %>%
  knitr::kable()
```




## Calculating relative entropy

The distance between (1) the race/ethnicity probability distribution for the US, ***Q***, and (2) the race/ethnicity probability distribution for some other geography, ***P***, is calculated below:
  
```{r message=FALSE, warning=FALSE}
gen_ent <- gen %>%
  filter(GEOID != '1' & !grepl('Puerto Rico', NAME)) %>%
  mutate(estimate = ifelse(estimate == 0, .0000001, estimate)) %>%
  left_join(usa, by = 'variable') %>%
  mutate(div = estimate * log(estimate/usa)) %>%
  group_by(GEOID, NAME) %>%
  summarize (relative_entropy = round(sum(div), 4),
             most_divergent = max(div), 
             most_divergent = variable[which(div == max(div))]) %>%
  ungroup() 
```



### States most similar to the US overall

The **most divergent** column specifies the sub-group that is most asymetrically distributed within a given geography  relative to the US overall. So, while the racial composition of Connecticut is very similar to the US overall, it differs most by being more White.


```{r message=FALSE, warning=FALSE}
gen_ent %>%
  filter(nchar(GEOID) == 2) %>%
  arrange(relative_entropy) %>%
  slice(1:10) %>%
  knitr::kable()
```



### Counties most similar to the US overall

```{r}
gen_ent %>%
  filter(nchar(GEOID) != 2) %>%
  arrange(relative_entropy) %>%
  slice(1:5) %>%
  knitr::kable()
```




### States least similar to the US overall

```{r message=FALSE, warning=FALSE}
gen_ent %>%
  filter(nchar(GEOID) == 2) %>%
  arrange(desc(relative_entropy)) %>%
  slice(1:10) %>%
  knitr::kable()
```




## Geographies

```{r message=FALSE, warning=FALSE, include=FALSE}
nonx <- c('78', '69', '66', '72', '60', '15', '02')

library(tigris)
options(tigris_use_cache = TRUE, 
        tigris_class = "sf")

us_counties <- tigris::counties(cb = TRUE) %>% 
  filter(!STATEFP %in% nonx)
us_states <- tigris::states(cb = TRUE) %>% 
  filter(!STATEFP %in% nonx)

laea <- sf::st_crs("+proj=laea +lat_0=30 +lon_0=-95")
us_counties <- sf::st_transform(us_counties, laea)
us_states <- sf::st_transform(us_states, laea)
```

```{r message=FALSE, warning=FALSE, eval=FALSE}
nonx <- c('78', '69', '66', '72', '60', '15', '02')

library(tigris)
options(tigris_use_cache = TRUE, 
        tigris_class = "sf")

us_counties <- tigris::counties(cb = TRUE) %>% 
  filter(!STATEFP %in% nonx)
us_states <- tigris::states(cb = TRUE) %>% 
  filter(!STATEFP %in% nonx)

laea <- sf::st_crs("+proj=laea +lat_0=30 +lon_0=-95")
us_counties <- sf::st_transform(us_counties, laea)
us_states <- sf::st_transform(us_states, laea)
```



### Relative entropy by county

```{r fig.height=5, message=FALSE, warning=FALSE}
us_counties %>%
  left_join(gen_ent, by = 'GEOID') %>%
  mutate(relative_entropy = ifelse(relative_entropy > 1, 
                                   1, relative_entropy)) %>%
  
  ggplot() + 
  geom_sf(aes(fill = relative_entropy),  lwd = 0) + 
  scale_fill_distiller(palette = "Blues", direction=1)+
  geom_sf(data=us_states, 
          fill = NA, 
          show.legend = F, 
          color="darkgray", 
          lwd=.5) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        legend.position = 'non') +
  labs(title = "Race-ethnicity relative entropy values by county")
```




### Most divergently distributed sub-group by county

```{r fig.height=5.5, message=FALSE, warning=FALSE}
us_counties %>%
  left_join(gen_ent, by = 'GEOID') %>%
  ggplot() + 
  geom_sf(aes(fill = most_divergent),  lwd = 0) + 
  ggthemes::scale_fill_stata() +
  geom_sf(data=us_states, 
          fill = NA, 
          show.legend = F, 
          color="darkgray", 
          lwd=.5) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        legend.position = 'bottom') +
  labs(title = "Most divergently distributed subgroup by county")
```


## References

Walker, K. (2018). Locating neighbourhood diversity in the American metropolis. *Urban Studies*, 55(1), 116-132.




