---
title: 'Roll call: 1965 Voting Rights Act'
date: '2020-08-11'
slug: roll-call-1965-voting-rights-act
---



```{r message=FALSE, warning=FALSE, include=FALSE}
notes_dir <- '/home/jtimm/jt_work/GitHub/technical_notes/resources/historic-cds/'
```

> Roll call/summary of the 1965 Voting Rights Act, passed by the House on 9 July 1965, and signed into law on 6 August 1965 by LBJ.  Here, a quick map of the House vote.


## Congressional districts in 1965

[PoliSci folks at UCLA](http://cdmaps.polisci.ucla.edu/) make available shape files for every U.S. Congressional District in use between 1789 and 2012.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tigris); options(tigris_use_cache = TRUE, tigris_class = "sf")

fname <- 'districts089'
nonx <- c('78', '69', '66', '72', '60', '15', '02')

states <- tigris::states(cb = TRUE) %>% 
  data.frame() %>%
  select(STATEFP, STUSPS) %>%
  rename(state_abbrev = STUSPS)

sf1 <- sf::st_read(dsn = paste0(notes_dir, fname), 
                        layer = fname, 
                        quiet = TRUE) %>%
  mutate(STATEFP = substr(ID, 2, 3),
         district_code = as.numeric(substr(ID, 11, 12))) %>%
  left_join(states) %>%
  filter(!STATEFP %in% nonx) %>%
  select(STATEFP, state_abbrev, district_code) 
```


```{r message=FALSE, warning=FALSE}
sf2 <- sf1 %>%
  rmapshaper::ms_filter_islands() %>% # not sure that this is doing 3-- 
  sf::st_transform(crs = "+init=epsg:4326") %>%
  sf::st_simplify(preserveTopology = TRUE, dTolerance = 0.005)
```



## VoteView: roll call & lawmaker details

Roll call for VOTING RIGHTS ACT OF 1965, and legislator details, eg, party affiliation, via `Rvoteview`.

```{r message=FALSE, warning=FALSE}
makers <- Rvoteview::download_metadata(type = 'members', 
                                       chamber = 'house', 
                                       congress = 89) %>%
  filter(state_abbrev != 'USA') %>%
  select(icpsr, state_abbrev, district_code, bioname, party_code) %>%
  mutate(party_code = ifelse(party_code==200, 'Rep', 'Dem'),
         icpsr = as.character(icpsr),
         district_code = as.integer(district_code)) 
```



```{r eval=TRUE, message=FALSE, warning=FALSE, include=FALSE}
vra <- Rvoteview::voteview_search('("VOTING RIGHTS ACT OF 1965") AND (congress:89) 
                                  AND (chamber:house)') %>%
                                  filter( date == '1965-07-09')

votes <- Rvoteview::voteview_download(vra$id)
names(votes) <- gsub('\\.', '_', names(votes))
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
vra <- Rvoteview::voteview_search('("VOTING RIGHTS ACT OF 1965") AND (congress:89) 
                                  AND (chamber:house)') %>%
                                  filter( date == '1965-07-09')

votes <- Rvoteview::voteview_download(vra$id)
names(votes) <- gsub('\\.', '_', names(votes))
```


Identify at-large representatives to address labeling discrepancies.

```{r}
atls <- makers %>%
  filter(bioname %in% votes$legis_data$name) %>%
  group_by(state_abbrev) %>%
  mutate(n = length(district_code)) %>%
  filter(n == 1)
```



```{r message=FALSE, warning=FALSE}
full <- votes$votes_long %>%
  inner_join(makers) %>%
  janitor::clean_names()%>%
  filter(state_abbrev != 'USA') %>%
  mutate(avote = case_when(vote == 1 ~ 'Yea',
                           vote == 6 ~ 'Nay',
                           !vote %in% c(1,6) ~ 'Not voting'),
         Party_Member_Vote = paste0(party_code, ': ', avote),
         ## fix at-large -- 
         district_code = ifelse(state_abbrev %in% atls$state_abbrev,
                                0, district_code))
```



```{r echo=FALSE}
set.seed(789)
full %>%
  select(state_abbrev:avote) %>%
  sample_n(5) %>%
  knitr::kable()
```


## Roll Call

```{r}
full$Party_Member_Vote <- factor(full$Party_Member_Vote)
full$Party_Member_Vote <- 
  factor(full$Party_Member_Vote, 
         levels(full$Party_Member_Vote)[c(3,6,1,4,2,5)])

full %>% 
  group_by(Party_Member_Vote) %>%
  count() %>%
  knitr::kable()
```




```{r message=FALSE, warning=FALSE}
fo_leaf <- sf2 %>%
  left_join(full %>% 
              ## weirdness re at-large @ 1965 -- TWO REPS/ONE GEO -- 
              group_by(state_abbrev, district_code, party_code) %>%
              slice(1) %>%
              mutate(district_code = ifelse(district_code == 98, 
                                            0, district_code)) %>%
              ungroup() ) %>%
  na.omit() %>%
  sf::st_set_crs('+proj=longlat +datum=WGS84')
```




```{r message=FALSE, warning=FALSE}
pal <- leaflet::colorFactor(palette = wnomadds::voteview_pal, 
                   domain = fo_leaf$Party_Member_Vote)

fl <- fo_leaf %>%
        leaflet::leaflet() %>%
        leaflet::addProviderTiles("CartoDB.Positron") %>%
        leaflet::setView(lng = -98.35, lat = 39.5, zoom = 4) %>%
  
  #options = providerTileOptions (minZoom = 6, maxZoom = 10)) %>%
        leaflet::addPolygons(label = ~bioname,
                    fill = TRUE,
                    stroke = TRUE, 
                    weight=1,
                    fillOpacity = 1,
                    color="white",
                    fillColor = ~pal(Party_Member_Vote)) %>%
  
        leaflet::addLegend("topleft", 
                  pal = pal, 
                  values = ~ Party_Member_Vote,
                  title = "Roll call",
                  opacity = 1)

widgetframe::frameWidget(fl, width="100%", height = '700')
```


<br>

## References

Jeffrey B. Lewis, Brandon DeVine, Lincoln Pitcher, and Kenneth C. Martis. (2013) Digital Boundary Definitions of United States Congressional Districts, 1789-2012. [Data file and code book]. Retrieved from http://cdmaps.polisci.ucla.edu on [10 August 2020].