---
title: 'roll call: two impeachments'
date: '2020-02-06'
slug: two-impeachments
---


Fifteen US Senators served in the impeachment trials of both President Clinton in 1999 and President Trump in 2020.  Here we take a quick look at how they voted using the `Rvoteview` package.  

The table below summarizes impeachment roll calls for the two 1999 articles & the two 2020 articles.  There was also a resolution to censure President Clinton.  (Note that Mitt Romney (UT) voted in support of Article I and against Article II in the 2020 impeachment trial of President Trump.) 

```{r message=FALSE, warning=FALSE}
library(tidyverse)

res <- Rvoteview::voteview_search("impeachment") %>%
  filter(chamber == 'Senate' & 
           date %in% c('2020-02-05', '1999-02-12')) %>%
  select(id, date, bill_number, text, question, yea, nay) %>%
  arrange(date) 
```

```{r}
res %>%
  select(-id) %>%
  mutate(article = c('-', 'I', 'II', 'II', 'I')) %>%
  arrange(date, article)%>% 
  knitr::kable()
```




Here we look at how the 15 US Senators -- members of both the 106th & 116th congresses -- voted on Article I from the 2020 trial and Article II of the 1999 trial.  

```{r message=FALSE, warning=FALSE, include=FALSE}
votes <- Rvoteview::voteview_download(res$id)
```


```{r eval=FALSE}
votes <- Rvoteview::voteview_download(res$id)
```


```{r message=FALSE, warning=FALSE}
sens <- Rvoteview:: member_search(chamber= 'Senate', 
                                  congress = c(106, 116))         

vs <- votes$votes.long %>% 
  inner_join(sens %>% 
               mutate(icpsr = as.character(icpsr))) %>%
  filter(vname %in% c('RS1060018', 'RS1160461')) %>%
  select(bioname, state_abbrev, 
         congress, party_name, vote) %>%
  mutate(vote = ifelse(vote == 1, 'Yea', 'Nay')) %>%
  spread(congress, vote) %>%
  filter(complete.cases(.)) %>%
  arrange(party_name, `106`) 

vs %>%
  knitr::kable()
```




So, seven senators voted both to **impeach President Clinton** and **acquit President Trump**.  Republicans all.


```{r}
vs1 <- vs %>% 
  filter(`106` ==  'Yea' & `116` == 'Nay') %>%
  mutate(bioname = gsub(',.*$', '', bioname),
         bioname = paste0(bioname,' (', state_abbrev, ')'))
vs1$bioname
```



