---
title: a relevant correspondance
date: '2020-01-23'
slug: a-letter-from-hamilton
---

An excerpt from [this letter](https://founders.archives.gov/documents/Hamilton/01-12-02-0184-0002) sent by Alexander Hamilton to George Washington on 1792-08-18, and quoted by Representative Adam Schiff on 2020-1-22 during the impeachment trial of Donald J. Trump.  Accessed here via [this Git Hub resource](https://github.com/jaytimm/founders_online_corpus) that makes the [Founders Online](https://founders.archives.gov/) database of writings/correspondances available as a collection of RDS files. 



```{r include=FALSE}
local <- '/home/jtimm/jt_work/GitHub/git_projects/FoundersArchiveCorpus/data/'
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
setwd(local)
gfiles <- list.files(path = local, 
                     pattern = "rds", 
                     recursive = TRUE) 

ffc_washington <- readRDS(gfiles[8])

ah <- ffc_washington %>% 
  filter(grepl('No popular Government was ever without its Catalines', og_text))
ah1 <- strsplit(ah$og_text, 'absurdity refutes itself.\n')[[1]][2]
ah2 <- strsplit(ah1, 'Objection the 15')[[1]][1]
ah3 <- gsub('When a man unprincipled', '**When a man unprincipled', ah2)
ah3 <- gsub('whirlwind', 'whirlwind**', ah3)
ah3 <- gsub('\n\n', '\n', ah3)
ah3 <- gsub('\n', '\n>\n', ah3)
ah3 <- paste('>', gsub(' *(\n*)*$', '\\1', ah3))
```



## Letter excerpt

Text in bold reflects portion quoted by Impeachment Manager Adam Schiff on 1/22/20.  Larger context (but not complete letter) presented below.

`r ah3`