---
title: vector space & homegrown analogies
date: '2020-07-05'
slug: vector-space-homegrown-analogies
---

A brief guide for building novel analogy questions for vector space model evaluation.  While most studies implement pre-canned analogy tasks, it is often useful to create, eg, content-specific analogy tasks.  Using the `text2vec` R-based NLP framework.


## Sample vector space model

For demonstration purposes, we (re-)use a model based on Google n-gram data (circa 1983-2008); available [here](https://github.com/jaytimm/google_ngrams_and_R/tree/master/google_one_percent_embeddings).  Model: PPMI-SVD; model details: window size = 5x5, dimensions = 250.


```{r include=FALSE}
library(tidyverse)
local_ngram <- '/home/jtimm/jt_work/GitHub/git_projects/google_ngrams_and_R/google_one_percent_embeddings'
git_files <- '/home/jtimm/jt_work/GitHub/technical_notes/resources/analogy/'
```


```{r message=FALSE, warning=FALSE}
setwd(local_ngram)
qvec <- readRDS('Q-[1983,2008].rds')
rownames(qvec) <- tolower(rownames(qvec))

# scale to 0-1 
qvec <- apply(qvec, 
              MARGIN = 2, 
              FUN = function(X) (X - min(X))/diff(range(X)))

str(qvec) # a numerical matrix -- 
```



## A set of semantic relations defined as `a : b`

Examples randomly assembled from web & randomly assigned categories for demonstration.

```{r}
new_ones <- c('hammer : nail',  'comb : hair',
              'white : black', 'up : down',
              'mansion : shack', 'yacht : dinghy',
              'short : light', 'long : heavy',
              'bees : hive',  'bears : den', 
              'speak : sing', 'walk : dance',
              'chef : food', 'sculptor : stone',
              'like : love', 'dislike : hate')

category <- c(rep('type1', 8), rep('type2', 8))

df <- data.frame(category, new_ones) %>%
  separate(new_ones, into = c('a', 'b'), sep = ' : ')

head(df) %>% knitr::kable()
```



## Build analogies 

As `a : b`' :: `a : b`" -- all permutations per group.

```{r}
build_questions <- function (x) { 
  
  lapply(unique(x$category), function(i) {
    
    comcaps <-subset(x, category == i)
    combos <- data.frame(t(combn(paste(comcaps$a, comcaps$b, 
                                       sep = ' '), m = 2))) 
    a <- paste(combos$X1, combos$X2, sep = ' ')
    b <- paste(combos$X2, combos$X1, sep = ' ')
    combos3 <- data.frame(analogy = c(a, b))
    combos3$category <- i
    combos3
    })
}
```




Apply function and format data as 'Google analogy' data structure

```{r}
new_test <- build_questions (df) %>% ###
  bind_rows() %>%
  group_by(category) %>%
  do(add_row(., .before=0)) %>%
  ungroup() %>%
  fill(category, .direction = c("up")) %>%
  mutate(category = paste0(': ', gsub(' |\\.txt$', '', category)),
         full = ifelse(is.na(analogy), category, analogy)) %>%
  select(full)
```




## Prepare analogies via `text2vec`

```{r}
setwd(git_files)
writeLines(new_test$full,'new_test.txt')
new_analogy_set <- text2vec::prepare_analogy_questions(
        questions_file_path = paste0(git_files, 'new_test.txt'), 
        vocab_terms = rownames(qvec))
```





## Evaluate via `text2vec`

```{r}
new_results <- text2vec::check_analogy_accuracy(
  questions_list = new_analogy_set, 
  m_word_vectors = qvec)
```



## Summary

For a slightly more detailed description of vector space evaluation via word analogy tasks, see this [brief note](https://www.puremoe.net/posts/vector-space-analogy-briefly/).


