---
title: counties, districts & republican vote shares
date: '2020-02-17'
slug: counties-districts-republican-vote-shares
---


Distribution of Republican vote by county & district for last three presidential elections.


## Counties

County-level voting returns for presidential elections 2008, 2012 & 2016. Via the [MIT Election Data and Science Lab](https://github.com/MEDSL/county-returns).

```{r message=FALSE, warning=FALSE}
library(tidyverse)
county <- read.csv(url('https://raw.githubusercontent.com/MEDSL/county-returns/master/countypres_2000-2016.csv')) %>%
  filter(party == 'republican' & year %in% c('2008', '2012', '2016')) %>%
  mutate(republican = round(candidatevotes/totalvotes*100,1), 
         geo = 'county') %>%
  select(geo, year, republican, totalvotes)
```


## Districts

Congressional district voting returns for the same elections.  Made available via `uspoliticalextras`. & from MIT Election Data and Science Lab.


```{r}
disty <- uspoliticalextras::uspol_dkos_returns_pres_cd %>%
  mutate(geo = 'district',
         totalvotes = 1) %>%
  select(geo, year, republican,  totalvotes)
```



## Republican vote shares

Percentage of total votes cast for the Republican nominee by geography. Counties are weighted by population, as approximated by total votes cast.  


```{r fig.height=6, message=FALSE, warning=FALSE}
bind_rows(disty, county) %>%
  group_by(geo, year) %>%
  mutate(weigh_group = sum(totalvotes)) %>%
  
  ggplot(aes(republican, color = geo, linetype = geo,
             weight = totalvotes/weigh_group)) + # weight = totalvotes,
  geom_density(size = 1) + #bw = 0.025
  facet_wrap(~year)
```


## References

McCarty, Nolan, Keith T Poole, and Howard Rosenthal. 2016. *Polarized America: The Dance of Ideology and Unequal Riches*. mit Press.


